% !TEX root = ../notes_template.tex
\chapter{Orthogonality}\label{chp:ortho}

% \minitoc

% Inset a quote for the chapter
\begin{flushright}
\textit{``Every pair of perpendicular vectors are orthogonal, but not every pair of orthogonal vectors is perpendicular.''}\\
% Draw a small line at the end of the quote
\rule{0.5\textwidth}{.4pt}\\
\textbf{Ben Grossmann} \small{in \href{https://math.stackexchange.com/questions/2326896/about-perpendicular-vectors}{StackExchange}.}
\end{flushright}

\section{Exercise}
\vspace{-0.5cm}
\begin{center}
    \rule{\textwidth}{1pt}
\end{center}

\begin{enumerate}
    \item Consider an orthonormal set of vectors,
    \[ V = \left\{ \mf{v}_1, \mf{v}_2, \ldots \mf{v}_r\right\} \quad \mf{v}_i \in \mb{R}^n \quad \forall i \in \left\{1, 2, \ldots r\right\}\]
    If there is a vector $\mf{w} \in \mb{R}^n$ such that $\mf{v}_i^T\mf{w} = 0\,\,\,\forall i \in \left\{1, 2, \ldots r\right\}$. Prove that $\mf{w} \notin span\left(V\right)$. 
    
    \item Prove that the rank of an orthogonal projection matrix $\mf{P}_{S} = \mf{UU}^T$ onto a subspace $\mc{S}$ is equal to the $\text{dim } \mc{S}$, where the columns of $\mf{U}$ form an orthonormal basis of $\mc{S}$. 
    
    \item For a matrix $\mf{A} \in \mb{R}^{m \times n}$, prove that $C\lp\mf{A}\rp \perp N\lp\mf{A}^T\rp$ and $C\lp\mf{A}^T\rp \perp N\lp\mf{A}\rp$. 

    \item If the columns of a matrix $\mf{A} \in \mb{R}^{n \times n}$ are orthonormal, prove that $\mf{A}^{-1} = \mf{A}^T$. What is $\mf{A}^T\mf{A}$ when $\mf{A}$ is rectangular $\left(\mf{A} \in \mb{R}^{m \times n}\right)$ with orthonormal columns? 

    \item What will happen when the Gram-Schmidt procedure is applied to: (a) orthonormal set of vectors; and (b) orthogonal set of vectors? If the set of vectors are columns of a matrix $\mf{A}$, then what are the corresponding $\mf{Q}$ and $\mf{R}$ matrices for the orthonormal and orthogonal cases? 
    
    \item If the columns of $\mf{A} \in \mb{R}^{m \times n}$ represent a basis for the subspace $\mc{S} \subset \mb{R}^m$. Find the orthogonal projection matrix $\mf{P}_\mc{S}$ onto the subspace $\mc{S}$. Hint: Gram-Schmidt orthogonalization. 

    \item Consider two orthogornal matrices $\mf{Q}_1$ and $\mf{Q}_2$. Is the $\mf{Q}_2^T\mf{Q}_1$ an orthogonal matrix? If yes, prove that it is so, else provide a counter-example showing $\mf{Q}_2^T\mf{Q}_1$ is not orthogonal. 

    \item Consider a 1 dimensional subspace spanned by the vector $\mf{u} \in \mb{R}^n$. What kind of a geometric operation does the matrix $\mf{I} - 2\frac{\mf{u}\mf{u}^T}{\mf{u}^T\mf{u}}$ represent? 

    \item Prove that when a triangular matrix is orthogonal, it is diagonal.
\end{enumerate}